### 2018-11-16 paper/spad_single/sections/introduction.tex
first draft of paper

contributions
%
\begin{itemize}
	\item introduce idea of RGB + single SPAD depth imaging		
	\item ... gesture recognition, face authentication, ... (e.g. ``learning to be a depth camera'' microsoft paper)
	\item build prototype, demonstrate results		
\end{itemize}