### 2018-11-16 paper/spad_single/sections/results.tex

\subsection{Implementation}

hardware, calibration, etc.

\subsection{Results}