### 2018-11-16 paper/spad_single/sections/method.tex
first draft of paper

\subsection{Monocular Depth Estimation}

\subsection{Global Depth Hints}

\begin{itemize}
	\item image formation model	
	\item training details
	\item simulations and comparison: qualitative and quantitative
\end{itemize}