### 2018-11-16 paper/spad_single/sections/related.tex

\paragraph{Depth Imaging}

\begin{itemize}
	\item stereo and multiview
	\item structured illumination and random patterns (kinect, etc.), active stereo
	\item time of flight (continuous wave and pulsed)
	\item what we do: like pulsed but much simpler setup; no scanning, no spad array, ...
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Monocular Depth Estimation}

\begin{itemize}
	\item summary of architectures and cost functions: u-net type architecture with reverse huber loss
	\item what we do: same thing, but augment with global hints (inspired by these approaches, we do ...)
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Deep Sensor Fusion}

global hints for super-resolution, colorization, depth estimation 
%
\begin{itemize}
	\item colorization
	\item david's 2018 paper for depth estimation and denoising (see david's 2019 sig paper for related work)
	\item what we do: slightly different application
\end{itemize}