### 2019-05-23 paper/spad_single/sections/results.tex
Working on the paper.

\subsection{Implementation}

hardware, calibration, etc.

\subsection{Results} 
% Table of quantitative results
\begin{table}
\begin{center}
\begin{tabular}{cccc|ccc}
  \toprule
    & $\delta^1 \uparrow$ & $\delta^2\uparrow$ & $\delta^3 \uparrow$ & RMSE $\downarrow$ & rel & $log_{10}$ \\
  \midrule
    DORN & 1 & 2 & 3 & 4 & 5 & 6 \\ 
  \bottomrule
\end{tabular} 
\end{center}
\end{table}

% Table of qualitative results
\begin{table}
  \begin{center}
    \begin{tabular}{cccccc}
      RGB & Albedo & Ground Truth & CNN & CNN + Raw Histogram & CNN + Albedo/Falloff \\
      \includegraphics[width=4cm]{sections/figures/comparison_table/rgb_0.png} &
      \includegraphics[width=4cm]{sections/figures/comparison_table/albedo_0.png} &
      \includegraphics[width=4cm]{sections/figures/comparison_table/gt_0.png} &
      \includegraphics[width=4cm]{sections/figures/comparison_table/cnn_0.png} &
      \includegraphics[width=4cm]{sections/figures/comparison_table/cnn_rawhist_0.png} &
      \includegraphics[width=4cm]{sections/figures/comparison_table/cnn_hist_0.png} \\ 
    \end{tabular}
  \end{center}
\end{table}