### 2019-08-08 paper/spad_single/sections/hardware.tex
Paper update

\subsection{Setup}
\begin{figure}
  \includegraphics[width=\textwidth/2]{sections/figures/hardware.jpeg}
  \caption{\textbf{Hardware setup}}
\end{figure}
\begin{itemize}
  \item{Description of hardware used}
  \item{Images of scenes used}
\end{itemize}
\begin{figure}
  \includegraphics[width=\textwidth/2]{sections/figures/hardware_results.jpeg}
  \caption{\textbf{Hardware results}}
\end{figure}