### 2019-08-08 paper/spad_single/sections/introduction.tex

contributions
%
\begin{itemize}
	\item We introduce the idea of augmenting an RGB camera with a single-pixel
    SPAD to address scale ambiguity error in monocular depth estimators.	
  \item We analyze our approach on indoor scenes using the NYU Depth v2 dataset.
    We demonstrate that our approach is able to resolve scale ambiguity while
    being fast and easy to implement.
	\item We build a hardware prototype and evaluate the efficacy of our
    approach on real-world data. 
\end{itemize}