### 2019-09-19 paper/spad_single/cvpr2019/supplement.tex
Finalizing captured results figure, adding results for densedepth.

\documentclass[10pt,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.
\usepackage{booktabs}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{array}
\usepackage{tabularx}
\usepackage{bm}
\usepackage{multirow}
\input{mark_defs.tex}
% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

%\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{4321}
\begin{document}

%%%%%%%%% TITLE
\title{Supplementary Information: Improving Monocular Depth Estimation with
  Global Depth Histogram Matching}

\maketitle
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_18_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_21_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_25_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_194_comparison.png}
  \caption{Results on DenseDepth}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_285_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_42_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_52_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_53_comparison.png}
  \caption{Results on DenseDepth}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_103_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_224_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_226_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_352_comparison.png}
  \caption{Results on DenseDepth}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_362_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_379_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_107_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_187_comparison.png}
  \caption{Results on DenseDepth}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_140_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_244_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_527_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_529_comparison.png}
  \caption{Results on DenseDepth}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_529_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_219_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_329_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_346_comparison.png}
  \caption{Results on DenseDepth}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_616_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_390_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_412_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_420_comparison.png}
  \caption{Results on DenseDepth}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_428_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_457_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_468_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/densedepth_477_comparison.png}
  \caption{Results on DenseDepth}
\end{figure*}
%   \includegraphics[height=3cm]{sections/figures/comparison/densedepth_499_comparison.png}
%   \includegraphics[height=3cm]{sections/figures/comparison/densedepth_258_comparison.png}
%   \includegraphics[height=3cm]{sections/figures/comparison/densedepth_341_comparison.png}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_8_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_14_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_15_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_23_comparison.png}
  \caption{Results on DORN}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_26_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_63_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_67_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_140_comparison.png}
  \caption{Results on DORN}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_170_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_522_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_548_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_569_comparison.png}
  \caption{Results on DORN}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_585_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_633_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_105_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_220_comparison.png}
  \caption{Results on DORN}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_252_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_259_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_178_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_219_comparison.png}
  \caption{Results on DORN}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_280_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_293_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_313_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_413_comparison.png}
  \caption{Results on DORN}
\end{figure*}
\begin{figure*}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_477_comparison.png}
  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_502_comparison.png}
%  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_428_comparison.png}
%  \includegraphics[width=\textwidth]{sections/figures/comparison/dorn_428_comparison.png}
  \caption{Results on DORN}
\end{figure*}
% Densedepth:
% - Office: 18, 21
% - Bathroom: 25, 194, 285
% - Bookstore: 42
% - Kitchen: 52, 53, 103, 224, 226, 352, 362, 379
% - Office: 107, 187
% - Living Room: 140, 244, 527, 529
% - Dining Room: 219, 329, 346, 616
% - Bedroom: 390, 412, 420, 428, 457*, 468*, 477, 499
% Failure Cases: 258, 338(?), 341, 537 649 (minor)
% 
% DORN:
% - Office: 8, 14, 15, 23, 26
% - Living room: 63, 67, 140, 170, 522, 548, 569, 585, 633
% - Office: 105, 220, 252, 259* (interesting)
% - Kitchen: 178
% - Dining Room: 219
% - Bathroom: 280, 293, 313
% - Bedroom: 413, 477, 502
% 
% Failure Cases: 96, 266(?)

\begin{figure*}[t!]
    \centering
    \begin{tabular}{p{5mm}*{4}{>{\centering\arraybackslash}p{1.15in}}c}
      \multirow[t]{5}{=}[-1in]{\rotatebox[origin=rc]{90}{Kitchen Scene}} & Ground Truth & CNN & CNN Mean Rescaled & CNN Histogram Matched & \\
      &
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_kitchen_scene/gt_z_proj_crop_depth_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_kitchen_scene/z_init_depth_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_kitchen_scene/z_med_scaled_depth_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_kitchen_scene/z_pred_depth_fig.png}&
      \includegraphics[height=1.15in]{sections/figures/captured/densedepth/8_29_kitchen_scene/depth_colorbar.pdf}\\
%      & RGB &  &  &  & \\
      &
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_kitchen_scene/rgb_cropped_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_kitchen_scene/z_init_diff_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_kitchen_scene/z_med_scaled_diff_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_kitchen_scene/z_pred_diff_fig.png}&
      \includegraphics[height=1.15in]{sections/figures/captured/densedepth/8_29_kitchen_scene/diff_colorbar.pdf}\\
      & RGB & \input{sections/figures/captured/densedepth/8_29_kitchen_scene/rmses.tex}& \\

      \rule{0pt}{3ex}  & & & & & \\
      \multirow[t]{3}{=}{\rotatebox[origin=c]{90}{Conference Room 1}}& 
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_conference_room_scene/gt_z_proj_crop_depth_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_conference_room_scene/z_init_depth_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_conference_room_scene/z_med_scaled_depth_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_conference_room_scene/z_pred_depth_fig.png}&
      \includegraphics[height=1.15in]{sections/figures/captured/densedepth/8_29_conference_room_scene/depth_colorbar.pdf}\\

      &
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_conference_room_scene/rgb_cropped_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_conference_room_scene/z_init_diff_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_conference_room_scene/z_med_scaled_diff_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_29_conference_room_scene/z_pred_diff_fig.png}&
      \includegraphics[height=1.15in]{sections/figures/captured/densedepth/8_29_conference_room_scene/diff_colorbar.pdf}\\
      & & \input{sections/figures/captured/densedepth/8_29_conference_room_scene/rmses.tex}& \\

      \rule{0pt}{3ex}  & & & & & \\
      \multirow[t]{3}{=}{\rotatebox[origin=c]{90}{Conference Room 2}}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_30_conference_room2_scene/gt_z_proj_crop_depth_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_30_conference_room2_scene/z_init_depth_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_30_conference_room2_scene/z_med_scaled_depth_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_30_conference_room2_scene/z_pred_depth_fig.png}&
      \includegraphics[height=1.15in]{sections/figures/captured/densedepth/8_30_conference_room2_scene/depth_colorbar.pdf}\\

      &
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_30_conference_room2_scene/rgb_cropped_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_30_conference_room2_scene/z_init_diff_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_30_conference_room2_scene/z_med_scaled_diff_fig.png}&
      \includegraphics[height=1.15in, width=1.15in]{sections/figures/captured/densedepth/8_30_conference_room2_scene/z_pred_diff_fig.png}&
      \includegraphics[height=1.15in]{sections/figures/captured/densedepth/8_30_conference_room2_scene/diff_colorbar.pdf}\\
      & & \input{sections/figures/captured/midas/8_30_conference_room2_scene/rmses.tex}& \\ 
    \end{tabular}
    \caption{Captured results initialized using the DenseDepth CNN.
      Second row shows absolute difference between above estimates and ground truth.}
    \label{fig:midas_captured}
%    \vspace{-1.5em}
\end{figure*}
\begin{figure*}[t!]
    \centering
    \begin{tabular}{p{5mm}*{4}{>{\centering\arraybackslash}p{1.15in}}c}
      \multirow[t]{5}{=}[-1in]{\rotatebox[origin=rc]{90}{Hallway}} & Ground Truth & CNN & CNN Mean Rescaled & CNN Histogram Matched & \\
      &
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_Hallway/gt_z_proj_crop_depth_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_Hallway/z_init_depth_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_Hallway/z_med_scaled_depth_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_Hallway/z_pred_depth_fig.png}&
      \includegraphics[height=1.15in]{sections/figures/captured/densedepth/8_30_Hallway/depth_colorbar.pdf}\\
      & & & & & \\

      & 
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_Hallway/rgb_cropped_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_Hallway/z_init_diff_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_Hallway/z_med_scaled_diff_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_Hallway/z_pred_diff_fig.png}&
      \includegraphics[height=1.15in]{sections/figures/captured/densedepth/8_30_Hallway/diff_colorbar.pdf}\\
      & RGB & \input{sections/figures/captured/densedepth/8_30_Hallway/rmses.tex}& \\

      \rule{0pt}{3ex}  & & & & & \\
      \multirow[t]{3}{=}{\rotatebox[origin=c]{90}{Poster}}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_poster_scene/gt_z_proj_crop_depth_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_poster_scene/z_init_depth_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_poster_scene/z_med_scaled_depth_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_poster_scene/z_pred_depth_fig.png}&
      \includegraphics[height=1.15in]{sections/figures/captured/densedepth/8_30_poster_scene/depth_colorbar.pdf}\\

      &
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_poster_scene/rgb_cropped_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_poster_scene/z_init_diff_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_poster_scene/z_med_scaled_diff_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_poster_scene/z_pred_diff_fig.png}&
      \includegraphics[height=1.15in]{sections/figures/captured/densedepth/8_30_poster_scene/diff_colorbar.pdf}\\
      & & \input{sections/figures/captured/densedepth/8_30_poster_scene/rmses.tex}& \\

      \rule{0pt}{3ex}  & & & & & \\
      \multirow[t]{3}{=}{\rotatebox[origin=c]{90}{Bookshelf}}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_small_lab_scene/gt_z_proj_crop_depth_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_small_lab_scene/z_init_depth_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_small_lab_scene/z_med_scaled_depth_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_small_lab_scene/z_pred_depth_fig.png}&
      \includegraphics[height=1.15in]{sections/figures/captured/densedepth/8_30_small_lab_scene/depth_colorbar.pdf}\\

      &
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_small_lab_scene/rgb_cropped_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_small_lab_scene/z_init_diff_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_small_lab_scene/z_med_scaled_diff_fig.png}&
      \includegraphics[width=1.15in, height=1.15in]{sections/figures/captured/densedepth/8_30_small_lab_scene/z_pred_diff_fig.png}&
      \includegraphics[height=1.15in]{sections/figures/captured/densedepth/8_30_small_lab_scene/diff_colorbar.pdf}\\
      & & \input{sections/figures/captured/densedepth/8_30_small_lab_scene/rmses.tex}\\

%      (a)&(b)&(c)&(d)&(e)\\
    \end{tabular}
    \caption{Captured results initialized using the DenseDepth CNN.
      Second row shows absolute difference between above estimates and ground truth.}
    \label{fig:midas_captured}
%    \vspace{-1.5em}
\end{figure*}

\end{document}