### 2019-10-03 paper/cvpr2020/main.tex

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.
%\usepackage{booktabs}
%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage{array}
%\usepackage{tabularx}
%\usepackage{bm}
%\usepackage{multirow}
%\usepackage{float}
%\usepackage[utf8x]{inputenc}

%\input{mark_defs.tex}

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\graphicspath{{"H:/Shared drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}{}{}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Optimizing Monocular Depth Estimation with Global Depth Histogram Matching using a Single SPAD}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
abstract
\end{abstract}

%%%%%%%%% BODY TEXT

\section{Introduction}
\label{sec:intro}
\input{sections/introduction}

\section{Related Work}
\label{sec:related}
\input{sections/related}

\section{Method}
\label{sec:method}
\input{sections/method}

\section{Evaluation and Assessment}
\label{sec:evaluation}
\input{sections/simulation}

\section{Experimental Demonstration}
\label{sec:prototype}
\input{sections/hardware}

\section{Discussion}
\label{sec:discussion}
\input{sections/discussion}

{\small
\bibliographystyle{ieee_fullname}
\bibliography{references}
}

\end{document}