### 2019-10-16 paper/cvpr2020/sections/method.tex
David's comments.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{method.pdf}
  \caption{\textbf{Overview of the full pipeline} We use a CNN to get an initial
  per-pixel depth estimate. Then we perform exact histogram matching using
  intensity-weighted pixel values on the corrected SPAD data.}
\end{figure}

Monocular depth estimation is a well-explored, yet active area of research (see Sec.~\ref{sec:related}). Any network trained for the MDE task takes as input an RGB image and estimates the depth map of the scene. Our approach augments existing MDE networks but is agnostic to the specific type of depth estimator. Indeed we show in Section~\ref{sec:evaluation} that our histogram matching technique equally improves a variety of different pre-trained estimators. In the remainder of this section, we therefore focus on modeling the image formation of a diffused single-photon avalanche diode (SPAD) and develop an approach to correcting an estimated depth map to match the global scene information captured by the SPAD.

%In this section, we describe the measurement model for a single-pixel time-of-flight lidar sensor under diffuse, pulsed laser illumination.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Image Formation Model of a Diffused SPAD}

Consider a diffused laser that emits a pulse at time $t = 0$ with time-varying
intensity $g(t)$ illuminating some 3D scene. We parameterize the geometry of the
scene as a distance map $z(x, y)$, where each of the 3D points has also some
unknown reflectivity $\alpha$ at the wavelength of the laser. Ignoring
interreflections of the emitted light within the scene, a single-pixel diffused
SPAD integrates the light scattered back from the scene towards the
detector as
%
\begin{equation}
	s \left( t \right)= \int_{\Omega_x} \int_{\Omega_y} \frac{\alpha \left( x,y \right)}{z(x,y)^2}  g \left( t - \frac{2z(x,y)}{c} \right) dx dy ,
	\label{eq:pulse_integral} 
\end{equation}  
%
where $c$ is the speed of light and $\Omega_{x,y}$ is the angular range of the
diffuser. In this formulation, we assume that the diffuser spreads light
uniformly over all angles. Each time such a light pulse is emitted into the
scene and scattered back to the detector,
the single-pixel SPAD time-stamps up to one of the returning photons with some
probability. The process is
repeated millions of times per second with the specific number of emitted pulses
being controlled by the repetition rate of the laser.
Each detected
photon arrival event is discretized into a histogram $h$ of the form
%
\begin{equation}
  h[n] \sim \mathcal{P} \left( \eta \int_{n\Delta t}^{(n+1)\Delta t} \left(f * s \right) \left( t \right)  dt + b \right),	
	\label{eq:spad_measurements}
\end{equation}
%
where $(n\Delta t, (n+1) \Delta t)$ models the $n^{th}$ time interval of the
temporal histogram, $\eta$ is the photon detection probability of the SPAD, $f$
is a function that models the temporal uncertainty in the detector, and $b$
represents ambient background light and false positive detections known as \textit{dark
count}. As derived in previous work, the combination of signal, background, and
noise can be modeled as an inhomogeneous Poisson process
$\mathcal{P}$~\cite{Kirmani:2014,Shin2015}. Finally, as in \cite{Xin2019}, we adopt the term
\textit{transient} for the histogram $h[n]$.

\subsection{Ambient Rejection and Falloff Correction}
Before histogram matching can be applied, the raw SPAD transient is processed in
three stages.

In the first stage, we remove the ambient and dark count photons from the transient. 
We compute an estimate of $b$ by assuming that the first $N$ bins of the SPAD transient
contain only ambient and dark count photons (this is equivalent to assuming the
minimum image distance is larger than $Nc\Delta T/2$) as:
\begin{equation}
  \hat b = \frac{1}{N}\sum_{n=0}^N h[n]. 
  \label{eq:ambient_estimate}
\end{equation}
We identify the first and last bins $n_{first}$ and $n_{last}$
corresponding to the minimum and maximum scene depths. We do this with a simple
edge thresholding procedure, computing the absolute first difference $d[n] =
\abs{h[n] - h[n+1]}$. If $h[n]$ and $h[n+1]$ are sampled according to $\mathcal{P}(b)$ and $b$ is large, then 
$h[n] - h[n+1]$ is well-approximated with a Gaussian with mean $0$ and variance $2b$.
Therefore, we can extract a candidate edge set $\mathcal{E}$ as:
\begin{equation}
  \mathcal{E} = \mset{n}{d[n] > \beta\sqrt{2\hat b}}.
  \label{eq:edge_set}
\end{equation}
In this work, we use $\beta = 5$, as it gives good results across both simulated
and captured data.
We compute initial estimates $n'_{first} = \min \mathcal{E}$ and
$n'_{last} = \paren*{\max \mathcal{E}} + 1$ respectively. By appealing to Proposition 3 in
\cite{Xin2019} and noting the similarity of our approach to collecting a single
non-line-of-sight SPAD measurement we can conclude that our estimates must be
close to the true values of $n_{first}$ and $n_{last}$.
We refine these estimates by walking them outwards until their bin values drop
below the ambient estimate $\hat b$, resulting in
$\hat n_{first} = \max\paren*{\set{n'_{first}}\cup\mset{n}{h[n] > \hat b, n \leq n'_{first}}}$ and
$\hat n_{last} = \min\paren*{\set{n'_{last}}\cup\mset{n}{h[n] > \hat b, n \geq n'_{last}}}$.
Finally, we reject all photons outside of the valid range by clamping those bins
to 0, and we subtract our ambient estimate from the photons within the valid
range:
\begin{equation}
  h'_{clean}[n] = \begin{cases}
    0 & n < \hat n_{first} \\
    \max(h[n] - \hat b, 0) & \hat n_{first} \leq  n \leq \hat n_{last} \\
    0 & n > \hat n_{last} \\
  \end{cases}.
  \label{eq:h_clean}
\end{equation}

In the second stage, we set 
\begin{equation}
  h_{clean}[n] = h'_{clean}[n] \cdot r_n^2
  \label{eq:h_scaled}
\end{equation}
where $r_n= \paren*{n + \frac12}\paren*{\frac{c\Delta t}{2}}$ is the depth
corresponding to bin $n$. This compensates for the radiometric falloff in a way
that is consistent with our hardware prototype's falloff model

In the final stage, we re-bin the transient, applying the Spacing-Increasing Discretization
(SID) of \cite{Fu2018} to produce $h_{target}$. This discretization scheme is
exponential rather than linear, allocating more bins to closer distances and fewer
bins to farther distances. The bin edges $t_i$ are given by the following equation,
parameterized by the number of bins $K$ and the range of distances $[\ell, u]$: 
\begin{equation}
  t_i = e^{\log(\ell) + \frac{\log(u/\ell) \cdot i}{K}}  \qquad i = 0,\ldots, K.
  \label{eq:sid_bin_edges}
\end{equation}
This reduces the number of bins in the histogram
matching procedure, giving speedups while maintaining performance. For the
simulated results,
we use $K = 140$ bins, and $(\ell, u) = (0.657, 9.972)$, (these numbers are
inspired by the discretization scheme for the original DORN network in
\cite{Fu2018})
 For the indoor caputred results, we use $K = 600$ bins, and $(\ell, u)=(0.4, 9) + \epsilon$ 
and for the outdoor captured results, we use $K=600$ and $(\ell, u)=(0.4, 11) +
\epsilon$ where $\epsilon \approx 0.094$ is the $z$ offset between the
SPAD and the Kinect (see Hardware for details).

% \begin{itemize}
%   \item Talk about histogram matching in the ideal case, jump straight to intensity 
%   \item Talk about histogram matching in our case, and how it approaches the
%     ideal case. Discuss the following corrections 
%     \begin{itemize}
%       \item Ambient/DC - Use \cite{Xin2019} to justify looking for large edges,
%         then the ambient estimate to get rid of the noise floor.
%       \item Falloff
%     \end{itemize}
%   \item Talk about how the histogram matching works with intensity
%     considerations applied, briefly.
%   \item We don't address jitter or poisson noise.
% \end{itemize}
% \begin{equation}
%   h[n] \sim \mathcal{P}\paren*{\sum_{x,y}\alpha_{x,y}\eta \lambda_{x,y}[n] + b} \label{global_hints}
% \end{equation}
% Given a SPAD with histogram $h$ according to the above equation, we first
% process the SPAD to remove the effects of some of the terms. First, we

%Neglecting albedo and falloff effects, an ideal detector counting photon events
%from a location $(x,y)$ in the time interval $(n\Delta t, (n+1) \Delta t)$ would record
%
%\begin{equation}
  %\lambda_{x,y}[n] = \int_{n\Delta t}^{(n+1) \Delta t} (f * g)\paren*{t - 2z(x,y)/c} dt \label{single_loc_spad} 
%\end{equation}  
%
%where $c$ is the speed of light, and $f$ is a function that models the temporal uncertainty in the
%detector. Single-photon avalanche diodes (SPADs) are highly sensitive
%photodetectors which are able to record single photon events with high temporal
%precision \cite{Stuff}. Since the event corresponding to the detection of a
%photon can be described with a Bernoulli random variable,
%the total number of accumulated photons in this time interval follows a Poisson
%distribution according to
%
%\begin{equation}
  %h[n] \sim \mathcal{P}\paren*{\sum_{x,y}\alpha_{x,y}\eta \lambda_{x,y}[n] + b} \label{global_hints}
%\end{equation}
%
%where $\alpha_{x,y} = r_{x,y}/z(x,y)^2$ captures the attenuation of the
%photon counts due to the reflectance $r(x,y)$ of the scene and due to the
%inverse square falloff $1/z(x,y)^2$.
%In addition, $\eta$ is the detection probability of a photon
%triggering a SPAD event, and $b = \eta a + d$ is the average number of background detections resulting
%from ambient photons $a$
%and erroneous ``dark count'' events $d$ resulting from noise within the SPAD.
%% \newpage
%% \begin{table*}[htbp]
%%   \begin{center}
  %%   \begin{tabularx}{\linewidth}{*{2}{X}}
  %%     \includegraphics[width=\textwidth/2-5pt]{sections/figures/spad_example/rgb.png} &
  %%     \includegraphics[width=\textwidth/2-5pt]{sections/figures/spad_example/rawdepth.png} \\
  %%     \includegraphics[width=\textwidth/2-5pt]{sections/figures/spad_example/depth_hist.png} &
  %%     \includegraphics[width=\textwidth/2-5pt]{sections/figures/spad_example/spad_hist.png} \\
  %%   \end{tabularx}
  %% \end{center}
  %% \caption{Sample Image. Top Left is the RGB image. Top Right is ground truth
  %%   depth. Bottom Left is Raw ground truth depth histogram. Bottom Right is
  %%   simulated SPAD measurements. Notice how closer depths are magnified and far
  %%   depths are attenuated.}
%% \end{table*}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Monocular depth estimation with global depth hints}
% Given a single RGB image $I(x,y)$ and a vector of photon arrivals $h[n]$
% described by equation \ref{global_hints}, we seek to
% reconstruct the ground truth depth map $z(x,y)$.
% Our method has two parts. First, we \textbf{initialize} our estimate of the depth map from the single RGB
% image via a monocular depth estimator described below. Second, we \textbf{refine} this depth map using
% the captured measurements $h[n]$ via exact histogram matching.

% \paragraph{Initialization}
% The first step in our method is to produce an initial estimate of ground truth
% depth. Convolutional Neural Networks have been shown to produce accurate, if poorly-scaled, estimates of depth
% from only a single image. We therefore choose to initialize our depth map
% estimate $\hat z^{(0)}(x,y)$ using
% a CNN. However, any depth estimator reliant on only a single
% view may be used for this step. Furthermore, in the larger context of our
% algorithm, it is more important that the network predict the correct ordinal
% relationships between pixels - that is, to predict the correct relative ordering
% of pixels $a$ and $b$, rather than to get all pixels exactly correct.

\subsection{Histogram Matching}
% \begin{algorithm}[H]
%  \caption{Exact Histogram Matching} 
%  \label{alg:ehm}
%  \input{sections/figures/exact_hist_matching_pseudo.tex}
% \end{algorithm}
Histogram matching is a procedure that takes a source image and a target
histogram and maps the pixel values of the source image so that the image's
histogram matches the target histogram. Given a single RGB image, we pass it
through an MDE to produce a depth map. In a parallel processing step,
we compute an estimate $W$ of the scene albedo using an intrinsic imaging method and
take the color channel of the resulting albedo image that corresponds to the
wavelength of the laser illumination.
We then compute a weighted histogram $h_{source}$ of the depth map $\hat z_0$ where the
per-pixel weights are given by $W$, and the bin edges are given by the SID
discretization algorithm in \cite{Fu2018}.

We match the histogram $h_{source}$ to $h_{target}$ using the method of
\cite{Morovic2002}. We compute a pixel movement
matrix $T$ such that $T[m, n]$ is the amount of mass that must be moved from $h_{source}[m]$
to $h_{target}[n]$. We do so using the procedure outlined in \ref{alg:ehm}

\begin{algorithm}[H]
 \caption{Pixel Movement} 
 \label{alg:ehm}
 \begin{algorithmic}
  \Procedure{FindMovement}{$h_s$ of length $M$, $h_t$ of length $N$} 
    \State Initialize $T$ as an $M \times N$ array of zeros.
    \For{$m$ in $1,\ldots,M$}
      \For{$n$ in $1,\ldots,N$}
        \State $p_s \gets \sum_{i=1}^{n-1} T[m, i]$
        \State $p_t \gets \sum_{i=1}^{m-1} T[i, n]$
        \State $T[m, n] \gets \min(h_s[m] - p_s, h_t[n] - p_t)$
      \EndFor
    \EndFor
    \State \textbf{return} $T$
  \EndProcedure
 \end{algorithmic}
\end{algorithm}

Finally, once we have obtained $T$, for a pixel $z_0[i, j]$ in bin $k$, we sample the new bin from the
distribution $T[k, :]/\sum_{n=1}^NT[k,n]$ (which is a distribution over target
depth bins $1,\ldots,N$) to get the bin of $\hat z[i,j]$. We can then apply the
standard SID decoding procedure to get the depth corresponding to that bin
\textcolor{red}{Does the above seem reasonable to include? Not sure if the
  pseudocode is helpful/necessary or not.}
%\subsection{Implementation Details}
%For the Monocular Depth Estimator, we use pretrained versions of the
%the Deep Ordinal Regression Network (DORN) \cite{} and the DenseDepth Network.
%The exact histogram matching method is as described in \cite{}.