### 2019-10-18 paper/cvpr2020/sections/discussion.tex

In summary, we demonstrate a method to greatly improve depth estimates from
monocular depth estimators by correcting the scale ambiguity errors inherent
with such techniques.  Our approach produces depth maps with accurate absolute
depth, and helps the generalization of neural networks for MDE across scene
types, including on data captured with our hardware prototype.  Moreover, we
require only minimal additional sensing hardware; we show that a single measurement
histogram from a diffused SPAD sensor contains enough information about global
scene geometry to correct errors in monocular depth estimates.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Limitations}

Since our method relies on the monocular depth estimate, the accuracy of the
output depth map also relies on the performance of the initial depth estimate.
The demonstrated results show that when the MDE technique produces a depth map
with good ordinal depth, where the ordering object depth is roughly correct, the
depth estimate can be corrected to produce accurate absolute depth. However,
errors in the ordinal depth may not necessarily be corrected in the histogram
matching procedure and so may propagate to the final output depth map.

Additionally, since our method relies on a diffused pulsed laser, the laser
power is spread out over the entire scene. So for distant scene points, very
little light will return to the SPAD, and it may be difficult to accurately
capture distant scene geometry in the histogram. Thus our method will likely be
best suited to short to medium-range scenes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Future Work}
While our hardware prototype is very large, future work could be done to
miniaturize this system. Our algorithm or similar sensor fusion algorithms could
also be integrated into electronics that already contain the required hardware
components, for example, existing cell phones with single-pixel SPAD
proximity sensors and RGB cameras.

Other methods for extracting scene information from the SPAD histogram could be
also employed, including learning-based methods to combine the MDE and
histogram. One might even consider sensing regimes where the number of
returning signal photons is very low, such as when the SPAD and camera operate
at high framerates. While most MDE techniques are tailored to clean RGB images,  
the SPAD histogram could potentially be used to help MDE
techniques generalize to noisy scenes under low-light conditions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Conclusions}
Neural network-based approaches to monocular depth estimation have yielded
significant improvements over the past several years for this task. However,
recently, such improvements have generally relied on new network architectures or
revised training procedures and have produced only modest improvements. Computational
photography approaches may provide the measurements needed to achieve much greater
performance improvements, especially as the sensors in devices we use every day
become more diverse, complex, and increasingly rely on techniques based on
sensor fusion and sophisticated computational algorithms.

%\begin{itemize}
%	\item new depth estimation CNNs achieve marginal gains, let's think about computational photography approaches that augment the measurements to achieve better performance
%	\item ...
%\end{itemize}