### 2019-10-18 paper/cvpr2020/sections/introduction.tex

Estimating dense 3D geometry from 2D images is an important open problem with applications to robotics, autonomous driving, and medical imaging. Depth maps are a common representation of scene geometry and are useful precursors to higher-level scene understanding tasks such as pose estimation and object detection. Additionally, many computer vision tasks rely on depth sensing, including navigation~\cite{geiger2013vision}, semantic segmentation~\cite{gupta2013perceptual,ren2012rgb,silberman2012indoor}, 3D object detection~\cite{gupta2014learning,lin2013holistic,shrivastava2013building,song2014sliding,song2016deep}, and 3D object classification~\cite{maturana2015voxnet,qi2016volumetric,wu20153d}.

Traditional depth sensing techniques include those based on stereo or multiview, active illumination, camera motion, or focus cues~\cite{szeliski2010computer}. However, each of these techniques has aspects that may make their deployment challenging. For example, stereo or multiview techniques require multiple cameras, active illumination techniques may have limited resolution or require time-consuming scanning procedures, and other techniques require camera motion or multiple exposures at different focus distances.

One of the most promising approaches to overcoming these challenges is monocular depth estimation (MDE), which requires only a single RGB image from a conventional camera to recover a dense depth map~\cite{Alhashim2018,Eigen2014,Fu2018,Laina2016,Saxena2006}. Recent approaches to MDE employ neural network models that learn to predict depth by exploiting monocular image cues such as perspective, occlusion, shading, and relative object size. While such models have significantly improved over recent years, MDE approaches to date are incapable of reliably estimating absolute distances in a scene due to the inherent scale ambiguities of monocular image cues. Instead, these models excel in predicting ordinal depth, or the relative ordering of objects in a scene~\cite{Eigen2014,Fu2018}. Interestingly, Alhashim and Wonka~\cite{Alhashim2018} recently showed that if the median ground truth depth of the scene is known, the initial output of a MDE network can be corrected to produce accurate absolute depth.

%For example, a larger object at a far distance to the camera results in the same 2D camera image as a smaller object of the same type that is closer to the camera. Therefore, MDE approaches to date are incapable of reliably estimating absolute distances of a scene. Interestingly, Alhashim and Wonka~\cite{Alhashim2018} recently showed that if the MDE network has oracle access to the ground truth median depth, then correcting the output of the CNN to match this median depth produces accurate absolute depth maps both qualitatively and quantitatively.

% move to discussion
%However, traditional approaches to depth estimation, such as stereo, suffer from lower performance when confronted with small angles or faraway objects. More exotic approaches use FMCW or time-of-flight LiDAR technologies, but these approaches are currently expensive and bulky.

%The most promising solution to these issues uses deep learning and convolutional neural networks to perform \textit{monocular depth estimation}, estimating dense depth maps from single RGB images.  However, this problem is underconstrained due to \textit{inherent scale ambiguity}, the unresolvable tradeoff between size and distance in single images. In practice, this issue commonly manifests itself in many monocular depth networks, and indeed, Wonka et. al. (cite) showed that if the method has oracle access to the ground truth median depth, then correcting the output of the CNN to match this median depth produces better depth maps both qualitatively and quantitatively.

\begin{figure}
  \includegraphics[width=\linewidth]{teaser.pdf}
  \caption{Monocular depth estimation predicts a depth map (upper right) from a single RGB image (upper left). The ill-posedness of the problem prevents reliable absolute depth estimation, resulting in large errors (inset image). The proposed method uses a measurement from a diffused SPAD (lower left), which resembles the shape of a histogram of the ground truth depth map, to correct the output of the depth estimation and optimize the quality of the estimated absolute depth (lower right).}
  \label{fig:teaser}
\end{figure}

Although access to the median ground truth depth is impossible in a realistic scenario, low-cost sensors capable of capturing aggregated depth information from a scene are readily available. For example, the proximity sensor on recent generation Apple iPhones uses a low-power pulsed light source and a single-pixel, single-photon avalanche diode (SPAD) to sense distance to an object directly in front of the phone. SPADs have also been used to recover full 3D geometry, forming the backbone of emerging single-photon LiDAR systems~\cite{Kirmani:2014,Li:2019,pawlikowska2017single}. However, SPADs have not been used for 3D imaging on consumer electronics because the requirement for ultra-fast timing electronics makes it difficult to produce high-resolution arrays at low cost, and because the scanning requirement for single-pixel systems introduces a point of mechanical failure and complicates high-resolution, high-framerate imaging.

%Individual SPADs are easy to fabricate and sufficiently cost-effective to be deployed in consumer electronics. When coupled with high-power pulsed lasers, they also form the backbone of emerging single-photon LiDAR systems~\cite{Kirmani:2014,pawlikowska2017single,Li:2019}. However, the required ultra-fast time-stamping electronics make it difficult to fabricate high-resolution SPAD arrays at low cost. Moreover, the scanning requirement for single-pixel systems precludes high-resolution, high-framerate depth mapping and is a barrier to integration in small consumer electronics.

Here, we propose to use a single-pixel SPAD and pulsed light source in an unconventional way: rather than optically focusing them to record the distance to a single scene point, we diffuse the emitted light and the detector over the entire scene. The resulting measurement resembles a histogram of the depth of the scene and can be used to achieve accurate absolute depth when combined with a monocular depth estimate (\cf~Figure~\ref{fig:teaser}).

%Technically, a diffused SPAD does not record a true depth histogram, but a transient that is affected by square-distance falloff of the backscattered light, spatially varying reflectance of the scene, ambient photons, and detector noise. Nevertheless, we can interpret the measurements of a single-pixel SPAD as something similar to a depth histogram.

%In this paper, we go further and show that by augmenting the RGB image with a histogram of global image depths, we can achieve substantially improved performance (and generalizability) over state-of-the-art monocular depth estimators. By performing an exact, weighted histogram matching on the output depth map of the depth estimator, we can match the depth histogram of the scene to the depth histogram of our estimate. This histogram matching is described in (cite) and is flexible enough to accommodate different pixel reflectances in the RGB image. Finally, this histogram can be captured relatively inexpensively using only a single pixel single-photon avalanche diode (SPAD) and pulsed laser illumination diffused over the field of view, representing a significant improvement in cost and simplicity over multi-pixel LiDAR arrays with expensive scanning mechanisms. It is worth noting that SPADs of this type have already made their way into existing smartphones, such as the iPhone X, and will likely play a role in future mobile sensing platforms as well.

% move to discussion
%Our method is not without limitations. It still requires a laser and single-pixel LiDAR detector, and as such, is sensitive to ambient photons. Being  a variant of histogram matching, our method is unable to transpose the values of pixels (i.e. if pixel $a$ is farther than pixel $b$ in the input, it will be farther than pixel $b$ in the output). In other words, our method is not able to resolve ordinal depth errors (errors where an object is wrongly placed closer or farther relative to another object). Finally, our method is non-differentiable, and is therefore unsuitable for end-to-end optimization of multi-part networks.

To this end, we develop a sensor fusion strategy that processes the ordinal depth computed by a monocular depth estimator to be consistent with the measurements captured by the diffused SPAD. We demonstrate in extensive simulations that our approach achieves substantial improvements in the quality of the estimated depth maps, regardless of which specific depth estimator is used. Moreover, we build an RGB-SPAD camera prototype and demonstrate in practice that our sensor fusion strategy achieves significantly higher-quality depth maps for a variety of captured scenes compared to MDE alone. With this work, we present a practical way to disambiguate depth estimation with RGB images using minimal additional sensing hardware. Specifically, we make the following contributions:
%
\begin{itemize}
	\item We introduce the idea of augmenting an RGB camera with a global depth histogram captured by a SPAD to address scale ambiguity error in monocular depth estimators.	
  \item We analyze this approach on indoor scenes using the NYU Depth v2 dataset and demonstrate that our approach is able to resolve scale ambiguity while being fast and easy to implement.
	\item We build a prototype RGB-SPAD camera and evaluate its efficacy on captured data, assessing both the quality and the ability of our method to help generalization of monocular depth estimators across scene types. 
\end{itemize}