### 2019-10-19 paper/cvpr2020/supplement.tex
Mark's edits.

\documentclass[10pt,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.
\usepackage{booktabs}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{array}
\usepackage{tabularx}
\usepackage{bm}
\usepackage{multirow}
\usepackage{float}
\usepackage[utf8x]{inputenc}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{algorithmicx}

\usepackage{dblfloatfix}

\input{mark_defs.tex}

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

%\cvprfinalcopy % *** Uncomment this line for the final submission

\graphicspath{{"figures/"}{"G:/Shared drives/Stanford Computational
    Imaging/Projects/single_spad_depth/figures/"}{"H:/Shared drives/Stanford Computational
    Imaging/Projects/single_spad_depth/figures/"}{"/Users/gordon/mount/teamdrive/Shared
    drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}
  {"/Volumes/GoogleDrive/Shared drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}}

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{4321}
\begin{document}

%%%%%%%%% TITLE
\title{Supplementary Information: Improving Monocular Depth Estimation with
  Global Depth Histogram Matching}

\maketitle
\section{Ablation study on number of SID bins}
% \begin{itemize}
%   \item Simulation: 70, 140, 210, 280 bins
%   \item Captured: 70, 140, 210, 280 bins
% \end{itemize}
% Compare RMSE and runtime.
We conducted an ablation study to assess the effect of the number of SID bins
\cite{Fu2018} on both runtime and RMSE. We performed this analysis using SPAD
data with a signal-to-background (SBR) of 100, simulated on the test set of NYU Depth v2. Our
MDE was DenseDepth, and only the actual histogram matching portion 
was timed, not the running of the MDE nor the denoising pipeline.
\begin{figure}[H]
  \centering
  \begin{tabular}{c|cc}
    \toprule
    \# of sid bins & RMSE & Approx. Time/image (sec) \\
    \midrule
    70  & $0.351$ & $0.24$ \\
    140 & $0.346$ & $0.63$ \\
    210 & $0.345$ & $1.12$ \\
    280 & $0.345$ & $1.84$ \\
    \bottomrule
  \end{tabular}
  \caption{Effect of number of SID bins on RMSE and runtime. The marginal
    improvement in RMSE is offset by the increase in runtime as the number of
    bins grows.}
  \label{fig:sid_ablation}
\end{figure}

\section{Ablation study on effect of reflectance estimation}
We conducted an ablation study to assess whether the use of a reflectance
estimate has an impact on the runtime and quality of the solution.
We performed this analysis using SPAD
data with a signal-to-background (SBR) of 100, simulated on the test set of NYU Depth v2. Our
MDE was DenseDepth, and only the actual histogram matching portion 
was timed, not the running of the MDE nor the denoising pipeline.

% \begin{itemize}
%   \item Intensity used to initialize histogram, not used to move pixels
%   \item Intensity used to initialize histogram and move pixels
%   \item Intensity not used to initialize histogram, used to move pixels
%   \item Intensity not used to initialize histogram, not used to move pixels
% \end{itemize}
% Compare RMSE and runtime.
\begin{figure}[H]
  \centering
  \begin{tabular}{c|c|cc}
    \toprule
    Intensity-weighted histogram & Intensity-aware pixel movement & Avg. RMSE & Approx. Time/image (sec) \\
    \midrule
    \multirow{2}{*}{Yes} & Yes & $0.346$ & $4.6$ \\
                         & No & $0.346$ & $0.6$ \\
    \midrule
    \multirow{2}{*}{No} & Yes & $0.444$ & $4.7$ \\
                        & No & $0.444$ & $0.6$ \\
    \bottomrule
  \end{tabular}
  \caption{Ablation study on simulated data for NYU Depth v2. When the SPAD
    is simulated with the reflectance info but no reflectance estimate is used
    to generate a weighted histogram from the CNN depth map, the results are
    significantly worse. Furthermore, once the pixel movement matrix has been
    computed, the pixel movement procedure need not take
  into account the weights of the pixels being moved, since doing so provides no
improvement and can take appreciably longer than a vectorized implementation
that does not take pixel weights into account.}
\end{figure}

\section{Additional results on NYU Depth v2}
Figures \ref{fig:densedepth_1}--\ref{fig:dorn_3} show additional results
for our method on the NYU Depth v2 dataset when the depth estimate is
initialized with the DenseDepth \cite{Alhashim2018} (Figures
\ref{fig:densedepth_1}--\ref{fig:densedepth_3}), and DORN \cite{Fu2018} (Figures
\ref{fig:dorn_1}--\ref{fig:dorn_3}) monocular depth estimators.
% and MiDaS \cite{MiDaS} (Figures \ref{midas_1} -- \ref{midas_3}).
% network of \cite{Alhashim2019}.
We compare the output of the network $z_0$, the
median-rescaled network output (where the depth map $z_0$ is scaled pixel-wise by a
scalar $\frac{\text{median}(z_{GT})}{\text{median}(z_0)}$, $z_{GT}$ being the
ground truth depth map), the network output matched to the ground truth depth histogram, and the output of
our histogram matching method under a signal-to-background ratio (SBR) of 100.
We use the Y channel of the YCbCr-transformed RGB image as our reflectance map
for both SPAD simulation and histogram matching.
We show absolute difference maps and also give
the root-mean-square error (RMSE) for each example.
\input{sections/supplement/simulation_comparison.tex}

\section{Additional results for hardware prototype}
Figures \ref{fig:midas_captured_1}--\ref{fig:dorn_captured_3} show all the
captured results when the depth estimate is initialized with the MiDaS
\cite{Lasinger:2019} (Figures \ref{fig:midas_captured_1}--\ref{fig:midas_captured_3}),
DenseDepth (Figures \ref{fig:densedepth_captured_1}--\ref{fig:densedepth_captured_3}),
and DORN (Figures \ref{fig:dorn_captured_1}--\ref{fig:dorn_captured_3}). We compare
the output of the network $z_0$, the mean-rescaled network output where the
depth map $z_0$ has been scaled pixel-wise by the scalar
$\frac{\text{median}(h_{target})}{\text{median}(z_0)}$ ($h_{target}$ is the
processed SPAD transient), and the output of our method. As our laser is red, we
use the R channel of the RGB image as our reflectance map. We show absolute
difference maps and also give the root-mean-square-error (RMSE) for each
example.

Black pixels in the ground truth depth correspond to locations where our scanner
was unable to produce an accurate depth estimate (this can occur for a variety
of reasons including dark albedo and surface specularity). These pixels are masked off and not used in the RMSE
calculation, and appear as an absolute difference of 0 in the difference maps.
\input{sections/supplement/hardware_comparison.tex}

{\small
\bibliographystyle{ieee_fullname}
\bibliography{references}
}

\end{document}