### 2020-02-29 paper/eccv2020/supplement.tex

% updated April 2002 by Antje Endemann
% Based on CVPR 07 and LNCS, with modifications by DAF, AZ and elle, 2008 and AA, 2010, and CC, 2011; TT, 2014; AAS, 2016; AAS, 2020

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{comment}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{color}

% INITIAL SUBMISSION - The following two lines are NOT commented
% CAMERA READY - Comment OUT the following two lines
\usepackage{ruler}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}

\usepackage{booktabs}
\usepackage{caption}
\usepackage{array}
\usepackage{tabularx}
\usepackage{bm}
\usepackage{multirow}
\usepackage{float}
\usepackage[utf8x]{inputenc}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{algorithmicx}
\usepackage{dblfloatfix}

\input{mark_defs.tex}
% \DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
% \def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}

\def\eg{\emph{e.g.}} \def\Eg{\emph{E.g.}}
\def\ie{\emph{i.e.}} \def\Ie{\emph{I.e.}}
\def\cf{\emph{c.f.}} \def\Cf{\emph{C.f.}}
\def\etc{\emph{etc}} \def\vs{\emph{vs.}}
\def\wrt{w.r.t.} \def\dof{d.o.f.}
\def\etal{\emph{et al.}}

\newcommand{\edit}[1]{\textcolor{red}{#1}}

\graphicspath{{figures/}{"G:/Shared drives/Stanford Computational
    Imaging/Projects/single_spad_depth/figures/"}{"H:/Shared drives/Stanford Computational
    Imaging/Projects/single_spad_depth/figures/"}{"/Users/gordon/mount/teamdrive/Shared
    drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}
  {"/Volumes/GoogleDrive/Shared drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}}
% Pages are numbered in submission mode, and unnumbered in camera-ready

\begin{document}
% \renewcommand\thelinenumber{\color[rgb]{0.2,0.5,0.8}\normalfont\sffamily\scriptsize\arabic{linenumber}\color[rgb]{0,0,0}}
% \renewcommand\makeLineNumber {\hss\thelinenumber\ \hspace{6mm} \rlap{\hskip\textwidth\ \hspace{6.5mm}\thelinenumber}}
% \linenumbers
\pagestyle{headings}
\mainmatter
\def\ECCVSubNumber{3668}  % Insert your submission number here

%\title{Improving Monocular Depth Estimation with Global Depth Histogram Matching using a Single SPAD Transient} % Replace with your title
\title{\textcolor{red}{Disambiguating Monocular Depth Estimation with a Single
    Transient: Supplemental Information}} % Replace with your title

% INITIAL SUBMISSION 
%\begin{comment}
\titlerunning{ECCV-20 submission ID \ECCVSubNumber} 
\authorrunning{ECCV-20 submission ID \ECCVSubNumber} 
\author{Anonymous ECCV submission}
\institute{Paper ID \ECCVSubNumber}
%\end{comment}
%******************

% CAMERA READY SUBMISSION
\begin{comment}
\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
\end{comment}
%******************
\maketitle

\section{Ablation study on number of SID bins}
% \begin{itemize}
%   \item Simulation: 70, 140, 210, 280 bins
%   \item Captured: 70, 140, 210, 280 bins
% \end{itemize}
% Compare RMSE and runtime.
We conducted an ablation study on the effect of the number of SID bins
\cite{Fu2018} on both runtime and RMSE. We performed this analysis using SPAD
data with a signal-to-background (SBR) of 100, simulated on the test set of NYU
Depth v2. We used DenseDepth~\cite{Alhashim2018} for our MDE CNN.
Only the histogram matching portion was timed, not the CNN nor the
denoising pipeline.

\begin{figure}[H]
  \centering
  \begin{tabular}{c|cc}
    \toprule
    \# of sid bins & RMSE & Approx. Time/image (sec) \\
    \midrule
    70  & $0.351$ & $0.24$ \\
    140 & $0.346$ & $0.63$ \\
    210 & $0.345$ & $1.12$ \\
    280 & $0.345$ & $1.84$ \\
    \bottomrule
  \end{tabular}
  \caption{Effect of number of SID bins on RMSE and runtime. The marginal
    improvement in RMSE is offset by the increase in runtime as the number of
    bins grows.}
  \label{fig:sid_ablation}
\end{figure}

\section{Ablation study on effect of reflectance estimation}
We conducted an ablation study on whether the use of a reflectance
estimate has an impact on the runtime and quality of the solution.
We performed this analysis using SPAD
data with a signal-to-background (SBR) of 100, simulated on the test set of NYU
Depth v2 and using DenseDepth~\cite{Alhashim2018} for our MDE CNN.
Only the histogram matching portion was timed, not the CNN nor the denoising pipeline.

% \begin{itemize}
%   \item Intensity used to initialize histogram, not used to move pixels
%   \item Intensity used to initialize histogram and move pixels
%   \item Intensity not used to initialize histogram, used to move pixels
%   \item Intensity not used to initialize histogram, not used to move pixels
% \end{itemize}
% Compare RMSE and runtime.
\begin{figure}[H]
  \centering
  \begin{tabular}{c|c|c|c}
    \toprule
    Intensity-weighted & Intensity-aware & Avg. & Time \\ histogram & pixel movement & RMSE & per image (sec) \\
    \midrule
    \multirow{2}{*}{Yes} & Yes & $0.346$ & $4.6$ \\
                         & No & $0.346$ & $0.6$ \\
    \midrule
    \multirow{2}{*}{No} & Yes & $0.444$ & $4.7$ \\
                        & No & $0.444$ & $0.6$ \\
    \bottomrule
  \end{tabular}
  \caption{Effect of reflectance modeling on RMSE and runtime. When the SPAD
    is simulated with the reflectance info but no reflectance estimate is used
    to generate a weighted histogram from the CNN depth map, the results are
    significantly worse. Furthermore, once the pixel movement matrix has been
    computed, the pixel movement procedure need not take
    into account the weights of the pixels being moved, since doing so provides no
    improvement and can take appreciably longer than a vectorized implementation
    that does not take pixel weights into account.}
\end{figure}

\section{Pseudocode, pixel shifting, and dither artifacts}
We give pseudocode for our algorithm here. In the first part of our algorithm we
compute the pixel shifting matrix mapping the histogram $h_s$ (computed from the
initial depth map and reflectance estimate)to $h_t$ (computed from the captured
transient).
{ \footnotesize
\begin{algorithm}
 \caption{Find Pixel Movement \edit{MOVE TO SUPPLEMENT!}} 
 \label{alg:ehm}
 \begin{algorithmic}
	\footnotesize
  \Procedure{FindPixelMovement}{$h_s$ of length $M$, $h_t$ of length $N$} 
    \State Initialize $T$ as an $M \times N$ array of zeros.
    \For{$m$ in $1,\ldots,M$}
      \For{$n$ in $1,\ldots,N$}
        \State $p_s \gets \sum_{i=1}^{n-1} T[m, i]$
        \State $p_t \gets \sum_{i=1}^{m-1} T[i, n]$
        \State $T[m, n] \gets \min(h_s[m] - p_s, h_t[n] - p_t)$
      \EndFor
    \EndFor
    \State \textbf{return} $T$
  \EndProcedure
 \end{algorithmic}
\end{algorithm}
}

Given this pixel movement matrix $T$, we apply the appropriate movements to the
initial depth map $I$. The pixels of the image $I$ take depth bin values in $\set{0,\ldots,K-1}$. 
{ \footnotesize
  \begin{algorithm}[H]
    \caption{Move Pixels}
    \label{alg:move_pixels}
    \begin{algorithmic}
      \footnotesize
      \Procedure{MovePixels}{input image $I$ size $M \times N$,
        pixel movement matrix $T$ of size $K \times K$}
      \For{$k$ in $0,\ldots, K-1$}
        \State $p[k,:] \gets T[k,:]/\sum_{i=1}^K T[k,i]$
      \EndFor
      \For{$m$ in $1,\ldots,M$}
        \For{$n$ in $1,\ldots,N$}
          \State Sample $k'$ according to $p[I[m,n],:]$.
          \State $I[m,n] \gets k'$.
        \EndFor
      \EndFor
      \State \textbf{return} $I$
      \EndProcedure
    \end{algorithmic}
  \end{algorithm}
} Because the pixel shifting process in Algorithm \ref{alg:move_pixels} contains
a sampling step, it is possible for
\textit{dither artifacts} to appear in the output image $I$, as shown in figure
\ref{fig:dither}. Specifically, when there are multiple possible output depth
bins for a given input depth bin, and a large region of equal depth in the input
image, the randomness in the pixel shifting algorithm will distribute the pixels
of large, equal-depth region in the input across the multiple possible output
depth bins in a random fashion.

\begin{figure}[H]
  \centering \includegraphics[width=0.66\textwidth]{dither.pdf}
  \caption{Example of dither artifacts. Sometimes, when our histogram matching
    is applied to images with large regions of similar depths, dither artifacts
    will occur.}
  \label{fig:dither}
\end{figure}

\section{Additional results on NYU Depth v2}
Figures \ref{fig:densedepth_1}--\ref{fig:midas_3} show additional results
for our method on the NYU Depth v2 dataset when the depth estimate is
initialized with the DenseDepth \cite{Alhashim2018} (Figures
\ref{fig:densedepth_1}--\ref{fig:densedepth_3}), DORN \cite{Fu2018} (Figures
\ref{fig:dorn_1}--\ref{fig:dorn_3}) 
and MiDaS \cite{Lasinger:2019} (Figures \ref{fig:midas_1} -- \ref{fig:midas_3}) monocular depth estimators.

We compare the output of the network $z_0$, the
median-rescaled network output (where the depth map $z_0$ is scaled pixel-wise by a
scalar $\frac{\text{median}(z_{GT})}{\text{median}(z_0)}$, $z_{GT}$ being the
ground truth depth map), the network output matched to the ground truth depth histogram, and the output of
our histogram matching method under a signal-to-background ratio (SBR) of 100.
We use the luminance of the RGB image as our reflectance map
for both SPAD simulation and histogram matching.
We show absolute difference maps and also give
the root-mean-square error (RMSE) for each example.
% \input{sections/supplement/simulation_comparison.tex}

\section{Additional results for hardware prototype}
Figures \ref{fig:midas_captured_1}--\ref{fig:dorn_captured_3} show all the
captured results when the depth estimate is initialized with the MiDaS
\cite{Lasinger:2019} (Figures \ref{fig:midas_captured_1}--\ref{fig:midas_captured_3}),
DenseDepth (Figures \ref{fig:densedepth_captured_1}--\ref{fig:densedepth_captured_3}),
and DORN (Figures \ref{fig:dorn_captured_1}--\ref{fig:dorn_captured_3}). We compare
the output of the network $z_0$, the mean-rescaled network output where the
depth map $z_0$ has been scaled pixel-wise by the scalar
$\frac{\text{median}(h_{target})}{\text{median}(z_0)}$ ($h_{target}$ is the
processed SPAD transient), and the output of our method. As our laser is red, we
use the R channel of the RGB image as our reflectance map. We show absolute
difference maps and also give the root-mean-square-error (RMSE) for each
example.

Black pixels in the ground truth depth correspond to locations where our scanner
was unable to produce an accurate depth estimate (this can occur for a variety
of reasons including dark albedo and surface specularity). These pixels are masked off and not used in the RMSE
calculation, and appear as an absolute difference of 0 in the difference maps.
% \input{sections/supplement/hardware_comparison.tex}

\clearpage

{\small
\bibliographystyle{splncs04}
\bibliography{references}
}
\end{document}