### 2020-07-19 paper/eccv2020CR/final/3668/3668.tex
Final camera-ready submission.

% updated April 2002 by Antje Endemann
% Based on CVPR 07 and LNCS, with modifications by DAF, AZ and elle, 2008 and AA, 2010, and CC, 2011; TT, 2014; AAS, 2016; AAS, 2020

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{comment}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{color}

% INITIAL SUBMISSION - The following two lines are NOT commented
% CAMERA READY - Comment OUT the following two lines
% \usepackage{ruler}
% \usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}

\usepackage{booktabs}
\usepackage{caption}
\usepackage{array}
\usepackage{tabularx}
\usepackage{bm}
\usepackage{multirow}
\usepackage{float}
\usepackage[utf8x]{inputenc}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{algorithmicx}
\usepackage{dblfloatfix}

\input{mark_defs.tex}
% \DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
% \def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}

\def\eg{\emph{e.g.}} \def\Eg{\emph{E.g.}}
\def\ie{\emph{i.e.}} \def\Ie{\emph{I.e.}}
\def\cf{\emph{c.f.}} \def\Cf{\emph{C.f.}}
\def\etc{\emph{etc}} \def\vs{\emph{vs.}}
\def\wrt{w.r.t.} \def\dof{d.o.f.}
\def\etal{\emph{et al.}}

\newcommand{\edit}[1]{\textcolor{black}{#1}}
% \newcommand{\edit}[1]{#1}

% \graphicspath{{figures/}{"G:/Shared drives/Stanford Computational
%     Imaging/Projects/single_spad_depth/figures/"}{"H:/Shared drives/Stanford Computational
%     Imaging/Projects/single_spad_depth/figures/"}{"/Users/gordon/mount/teamdrive/Shared
%     drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}
%   {"/Volumes/GoogleDrive/Shared drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}}
\graphicspath{{figures/}}
% Pages are numbered in submission mode, and unnumbered in camera-ready

\begin{document}
% \renewcommand\thelinenumber{\color[rgb]{0.2,0.5,0.8}\normalfont\sffamily\scriptsize\arabic{linenumber}\color[rgb]{0,0,0}}
% \renewcommand\makeLineNumber {\hss\thelinenumber\ \hspace{6mm} \rlap{\hskip\textwidth\ \hspace{6.5mm}\thelinenumber}}
% \linenumbers
\pagestyle{headings}
\mainmatter
\def\ECCVSubNumber{3668}  % Insert your submission number here

\title{Disambiguating Monocular Depth Estimation with a Single Transient} % Replace with your title

% CAMERA READY SUBMISSION
% \begin{comment}
\titlerunning{Disambiguating Monocular Depth Estimation with a Single Transient}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Mark Nishimura\orcidID{0000-0003-3976-254X}  \and
David B. Lindell\orcidID{0000-0002-6999-3958} \and
Christopher Metzler\orcidID{0000-0001-6827-7207} \and \\
Gordon Wetzstein\orcidID{0000-0002-9243-6885}}
%
\authorrunning{M. Nishimura et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Stanford University, Stanford, CA \\
\email{\{markn1, lindell, cmetzler, gordon.wetzstein\}@stanford.edu}}
%******************
\maketitle

\begin{abstract}
  Monocular depth estimation algorithms successfully predict the
  relative depth order of objects in a scene.
  However, because of the fundamental scale ambiguity associated with monocular
  images, these algorithms fail at correctly predicting true metric
  depth. In this work, we demonstrate how a depth histogram of the scene, which
  can be readily captured using a single-pixel time-resolved
    detector, can be fused with the output of existing monocular depth
  estimation algorithms to resolve the depth ambiguity problem.
  We validate this novel sensor fusion technique experimentally and in extensive
  simulation. We show that it significantly improves the performance of several
  state-of-the-art monocular depth estimation algorithms.  
  \keywords{depth estimation, time-of-flight imaging}
\end{abstract}

\vspace{-5pt}
\section{Introduction}
\vspace{-5pt}
\label{sec:intro}
\input{sections/introduction}

\vspace{-5pt}
\section{Related Work}
\vspace{-5pt}
\label{sec:related}
\input{sections/related}

\vspace{-5pt}
\section{Method}
\vspace{-5pt}
\label{sec:method}
\input{sections/method}

\vspace{-5pt}
\section{Evaluation and Assessment}
\vspace{-5pt}
\label{sec:evaluation}
\input{sections/simulation}

\vspace{-5pt}
\section{Experimental Demonstration}
\vspace{-5pt}
\label{sec:prototype}
\input{sections/hardware}

\vspace{-5pt}
\section{Discussion}
\vspace{-5pt}
\label{sec:discussion}
\input{sections/discussion}

\vspace{-4 pt}
\subsection*{Acknowledgments}
D.L.~was supported by a Stanford Graduate Fellowship. 
C.M.~was supported by an ORISE Intelligence Community Postdoctoral Fellowship. G.W.~was supported by an NSF CAREER Award (IIS 1553333), a Sloan Fellowship, by the KAUST Office of Sponsored Research through the Visual Computing Center CCF grant, and a PECASE by the ARL.

\clearpage
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
% \bibliography{references}
\input{3668.bbl}

\end{document}